EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-04-28 05_examples.tex
Updated draft. * added a section of parse_Umatrix (to be written) * drafting attractive HM, figure loaded.

\subsection{Bethe lattice DMFT (Fortran API)}
The description of the Mott or metal to insulator transition (MIT)
within the Bethe lattice, i.e. Cayley tree with connection
$z\to\infty$ is conventionally considered the test bed of any DMFT
application. Here  we discuss a guided implementation based on the
Fortran API of \NAME of the DMFT solution of the Bethe lattice at
$T=0$. 
We consider a Fermi-Hubbard model:
$$
H = -t \sum_{\langle ij\rangle\s}c^+_{i\s}c_{j\s} + U \sum_i n_{i\up}n_{i\dw}
$$
defined on a Bethe lattice with density of states
$\rho(x)=\frac{1}{2D}\sqrt{D^2-x^2}$, where $D=2t$ is the
half-bandwidth~\cite{Georges1996}.
In DMFT this model is mapped
onto a quantum impurity problem with an effective electronic bath
which needs to be determined self-consistently.
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK2
   USE SCIFOR
   implicit none
   !> Number of discretization points for the Bethe DOS
   integer,parameter                           :: Le=5000
   !> Bethe half-bandwidth = energy unit
   real(8),parameter                           :: D=1d0
   !>Bath:
   real(8),allocatable                         :: Bath(:)
   !> local dynamical functions, rank-5 [Nspin,Nspin,Norb,Norb,L]
   complex(8),allocatable,dimension(:,:,:,:,:) :: Weiss,Smats,Sreal,Gmats,Greal,Weiss_
   ... (omissis)
   !> READ THE input using EDIpack procedure:
   call ed_read_input(trim(finput))
   !
   !> Setup the Bethe DOS $\rho$ (linear dispersion)
   Ene = linspace(-D,D,Le,mesh=de)
   DOS = dens_bethe(Ene,D)
\end{lstlisting}

@@

\subsubsection{Results}
In the following we present some results obtained by executing this
simple program varying the interaction strenght :f:var:`uloc`.
Differently from the previous case of a quantum impurity embedded in a
given bath describing the progressive formation of a
strongly renormalized Fermi liquid state, here the DMFT
self-consistency allows to describe the transition from a correlated
metal to a Mott insulating state.

@@

\subsection{Attractive Hubbard model (Python API)}
The second application has the twofold aim to demonstrate the
superconductivity support of the \NAME library as well as to
illustrate the Python API through a working example.

>>>>

\subsection{Bethe lattice DMFT (Fortran API)}\label{SecExamplesBetheDMFT}
The description of the Mott or metal to insulator transition (MIT)
within the Bethe lattice, i.e. Cayley tree with connection
$z\to\infty$ is conventionally considered the test bed of any DMFT
application. Here  we discuss a guided implementation based on the
Fortran API of \NAME of the DMFT solution of the Bethe lattice at
$T=0$. 
We consider a Fermi-Hubbard model:
$$
H = -t \sum_{\langle ij\rangle\s}c^+_{i\s}c_{j\s} + U \sum_i n_{i\up}n_{i\dw}
$$
where $c^+_{i\s}$ ($c_{i\s}$) is the creation (destruction) operator for an electron at site $i$
and with spin $\s$, $n_{i\s}=c^+_{i\s}c_{i\s}$ is the occupation
operator and the first sum is over the nearest neighbor sites $\langle
ij\rangle$. 
The system is defined on a Bethe lattice with density of states
$\rho(x)=\frac{1}{2D}\sqrt{D^2-x^2}$, where $D=2t$ is the
half-bandwidth~\cite{Georges1996}.
In DMFT this model is mapped
onto a quantum impurity problem with an effective electronic bath
which needs to be determined self-consistently.
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK2
   USE SCIFOR
   implicit none
   !> Number of discretization points for the Bethe DOS
   integer,parameter                           :: Le=5000
   !> Bethe half-bandwidth = energy unit
   real(8),parameter                           :: D=1d0
   !>Bath:
   real(8),allocatable                         :: Bath(:)
   !> local dynamical functions, rank-5 [Nspin,Nspin,Norb,Norb,L]
   complex(8),allocatable,dimension(:,:,:,:,:) :: Weiss,Smats,Sreal,Gmats,Greal,Weiss_
   ... (omissis)
   !> READ THE input using EDIpack procedure:
   call ed_read_input(trim(finput))
   !
   !> Setup the Bethe DOS $\rho$ (linear dispersion)
   Ene = linspace(-D,D,Le,mesh=de)
   DOS = dens_bethe(Ene,D)
\end{lstlisting}

@@

\paragraph{Results}
In the following we present some results obtained by executing this
simple program varying the interaction strenght :f:var:`uloc`.
Differently from the previous case of a quantum impurity embedded in a
given bath describing the progressive formation of a
strongly renormalized Fermi liquid state, here the DMFT
self-consistency allows to describe the transition from a correlated
metal to a Mott insulating state.

@@

\subsection{Attractive Hubbard model (Python API)}\label{SecExamplesAHM}
The second application has the twofold aim to demonstrate the
superconductivity support of the \NAME library as well as to
illustrate the use of the Python API through a working example.
We consider the simple, yet non-trivial, case of the attractive
Hubbard model on a two-dimensional square lattice:

$$
H = \sum_{\ka\s} \e(\ka)c^+_{\ka\s}c_{\ka\s} -U\sum_i n_{i\up}n_{i\dw}
$$
where $U>0$, $c^+_{\ka\s}=\tfrac{1}{\sqrt{N}}\sum_i e^{-i\ka\cdot R_i} c^+_{i\s}$
and $c^+_{i\s}$ is the creation operator for an electron at site $i$
and with spin $\s$, $n_{i\s}=c^+_{i\s}c_{i\s}$ is the occupation
operator and $\e(\ka)=-2t[\cos{(k_x a)}+ \cos{(k_y a)}]$ is the
energy dispersion relation. In the following we the lattice distance
$a=1$  and the energy unit as $4t=D=1$.

The DMFT workflow for this case is very similar to the previous one
provided works in the Nambu basis defined by the spinor
$\psi_i=[c_{i\up}, c^+_{i\dw}]^t$.
In this basis any correlation functions, e.g. the Green's function, takes the matrix form:
\begin{equation}
  G =
  \begin{pmatrix}
    \hat{G}_{\uparrow\uparrow} & \hat{F}_{\uparrow\downarrow}\\
    \hat{\bar{F}}_{\downarrow\uparrow}  &    \hat{\bar{G}}_{\downarrow\downarrow} \\
   \end{pmatrix}
\end{equation}
where the symbol $\hat{A}$ indicates the possible multi-orbital
character which reduced to a scalar in the present case of a single
orbital. The symbol $\hat{\bar{A}}$ indicates that the components in
the second row are connect by particle-hole and time-reversal
symmetries to the first row one. The specific relation depends on the
symmetry of the order parameter (here $s$-wave) and the nature of the
complex argument of the functions, i.e. Matsubara or real-axis:
\begin{equation}
\begin{array}{cc}
  \hat{\bar{G}}(i\omega) = -\hat{G}^*(i\omega)\;; &  \hat{\bar{F}}(i\omega) = \hat{F}(i\omega)\\
  \hat{\bar{G}}(\omega)  = -\hat{G}^*(-\omega) \;; & \hat{\bar{F}}(\omega) = \hat{F}^*(i\omega)\\
\end{array}
\end{equation}  
The code implementation closely follows the previous structure, with
some notable exceptions related to the Nambu structure for which,
using the symmetries above, we only need to deal with the first row
independent components. In the first part of the code we initialize
the lattice structure and the solver:
\begin{lstlisting}[style=mypython,numbers=none,basicstyle={\scriptsize\ttfamily}]
import numpy as np
from edipy2 import global_env as ed
import mpi4py
from mpi4py import MPI
import os,sys
from aux_funx import * #include function to build $G_{loc}$ and Weiss fields

comm = MPI.COMM_WORLD
rank = comm.Get_rank()
master = (rank==0)

#Functions
def generate_kgrid(Nk):
    b1=2*np.pi*np.array([1.0,0.0])
    b2=2*np.pi*np.array([0.0,1.0])
    n1, n2 = np.meshgrid(np.arange(Nk), np.arange(Nk))
    n1=n1/Nk;n2=n2/Nk
    gridout = np.stack([n1.ravel(), n2.ravel()], axis=-1)
    return np.dot(gridout,[b1,b2])

def h_square2d(k,t):
  return -2*t*(np.cos(k[...,0,np.newaxis,np.newaxis])+np.cos(k[...,1,np.newaxis,np.newaxis]))*np.eye(ed.Norb)

ed.read_input("inputAHM.conf")

kgrid = generate_kgrid(Nk)

#Generate hk and hloc
Hk   = h_square2d(kgrid,t_hop)
HkNambu   = np.array([h_square2d(kgrid,t_hop),-np.conj(h_square2d(-kgrid,t_hop))])
Hloc = np.sum(Hk,axis=0)/Nk**2
Hloc = Hloc.astype(complex)
ed.set_hloc(Hloc)

#SETUP SOLVER
Nb=ed.get_bath_dimension()
bath = ed.init_solver()
bath_prev = np.copy(bath)
\end{lstlisting}

The iterative scheme for the solution of DMFT closely follows the
sequence discussed in \secu{SecExamplesBetheDMFT}:   
\begin{enumerate}
\item Call the exact diagonalization {\bf impurity solver} {\tt
    ed\_solve} providing the set of parameters $\vec{x}$ as input. 
\item Retrieve the self-energy functions
  $\hat{\Sigma}_{\a\b\up\up}(i\omega)$ and $\hat{S}_{\a\b\up\dw}(i\omega)$ on the 
  Matsubara axis using {\tt ed\_get\_sigma}.
\item
  Evaluate the interacting local Green's functions $\hat{G}_{loc}$ and
  $\hat{F}_{loc}$:
  \begin{equation}
  G_{loc}(i\omega) =
  \int_\RRR d\e \rho(\e)
  \begin{pmatrix}
    (i\omega +\mu)\hat{\11} - \hat{h}^0 - \hat{\Sigma}(i\omega) -\e & -\hat{S}(i\omega) \\
    -\hat{S}(i\omega) & (i\omega +\mu)\hat{\11} + \hat{h}^0 +
    \hat{\Sigma}^*(i\omega) +\e\\
  \end{pmatrix}^{-1}
\end{equation}
\item Update the normal and anomalous Weiss fields, respectively 
  $\GG^{-1}$ and $\Theta^{-1}$, through the {\bf self-con\-sis\-ten\-cy}
    relation: $\GG^{-1}_0(i\omega) = G_{loc}(i\omega) +
    \Sigma(i\omega)$ in Nambu space. 
  \item Optimize the bath parameters $\vec{x}$ to best describe the  updated
    Weiss fields. 
\end{enumerate}
which is implemented in Python as follows:
\begin{lstlisting}[style=mypython,numbers=none,basicstyle={\scriptsize\ttfamily}]
#DMFT CYCLE
converged=False;iloop=0
while (not converged and iloop<ed.Nloop ):
    iloop=iloop+1
    #Solve the EFFECTIVE IMPURITY PROBLEM (first w/ a guess for the bath)
    ed.solve(bath)
    print(ed.get_phi())

    #Use aux_funx procedures:
    Smats = np.array([ed.get_sigma(axis="m",typ="n"),ed.get_sigma(axis="m",typ="a")])    
    Gmats = get_gloc(wm*1j       ,ed.xmu,HkNambu,Smats,axis="m") 
    Weiss = dmft_weiss_field(Gmats,Smats)

    #Print...

    #Fit
    bath = ed.chi2_fitgf(Weiss[0],Weiss[1],bath)

    if(iloop>1):
        bath = wmixing*bath + (1.0-wmixing)*bath_prev
    bath_prev=np.copy(bath) 
    err,converged=ed.check_convergence(Weiss,ed.dmft_error)

ed.finalize_solver()
\end{lstlisting}

\paragraph{Results}
We showcase here some results for the attractive Hubbard model to
illustrate the capabilities of \NAME to solve $s$-wave superconducting
impurity problems at zero and finite temperatures. 
In figure...
\begin{figure}[t!]
  \includegraphics[width=\linewidth]{figures/figAHM.pdf}
    \caption{\label{fig2}%
      \textbf{The BCS to BEC crossover.}
      (a) Evolution of the spectral functions
      $-\Im{G_{loc}(\omega)}/\pi$ as a function of increasing
      attraction $U$. 
      (b) The order parameter $\phi=\langle c_\up c_\dw\rangle$ as a
      function of the attraction $U$. Data for BCS (gray) is compared
      to DMFT results (blue line and symbols). 
      (c) The correlation strength
      $\Theta=\frac{|S(0)-S(\infty)|}{S(\infty)}$ as a function of the
      attraction $U$ across the BCS-BEC crossover. 
      (d) Comparison of the double occupation $D=\langle n_\up
      n_\dw\rangle$ as a function of attraction $U$ between BCS (gray
      solid line) and DMFT (blue solid line and symbols). 
      (e) Superconducting order
      parameter $\phi$ as a function of temperature across the
      superconductor-to-normal phase transition. Data for $U=4$. The
      fit highlights the critical behavior with a mean-field exponent
      $\beta=1/2$ (red dashed line).       
        }
\end{figure}
